---
title: "Commenter COI Extraction Pipeline"
author: Spencer Kuhn
output:
  pdf_document: default
bibliography: "../../Dropbox/Literature Reviews/references.bib"
link-citations: yes
---

```{r setup}

# reset global environment ----
rm(list = ls())

# import packages ----
library(dplyr)
library(data.table)
library(lubridate)
library(ellmer)
library(sentimentr)

# TODO: include fuzzy matching with agrap package

```

The GeoNames database compiles place names alongside their corresponding point or polygon geometries at all administrative levels around the world [@GeoNames]. Pennsylvania GeoNames locations serve as the building block unit of each commenter's community of interest.

```{r geonames-data-import}

# import pennsylvania geonames ----
pennsylvaniaGeoNames <- readRDS(file = "./Data/PennsylvaniaGeoNames.rds")

# import us geonames (source data too large for github) ----
# pennsylvaniaGeoNames <- read.delim("./Data/US/US.txt", header = FALSE) |>
#   stats::setNames(
#     nm = c(
#       "Geonameid", "Name", "Asciiname", "AlternateNames",
#       "Latitude", "Longitude",
#       "FeatureClass", "FeatureCode",
#       "CountryCode", "CC2",
#       "Admin1Code", "Admin2Code", "Admin3Code", "Admin4Code",
#       "Population", "Elevation", "Dem", "Timezone",
#       "ModificationDate"
#     )
#   ) |>
#   dplyr::filter(CountryCode == "US", Admin1Code == "PA") |>
#   dplyr::select(-c(CC2, Admin4Code, Dem, Timezone, ModificationDate))

```

Public testimony data was gathered from the Redistricting Data Hub, which has compiled comments from the 2020 redistricting cycle for 14 individual states [@RedistrictingDataHubPublicTestimony]. Tabula, a software capable of converting PDFs to csv files, was used to process comments about Pennsylvania state house boundaries and commenter communities of interest.

```{r pa-commenter-data-import}

# import tabula table for pennsylvania house redistricting comments ----
tabulaPAHouse <- read.csv(file = "./Data/TabulaPAHouse.csv") |>
  dplyr::select(-X) |>
  dplyr::rename(CommunityName = "Community.Name") |>
  dplyr::mutate(
    Description = paste0(CommunityName, ": ", Description),
    Date = lubridate::dmy(x = Date)
  )

```

The ellmer package in R was used to interact with off-the-shelf LLM chat features and extract structured data [@ellmer], including comment sentiment, comment clarity, the locations the commenter mentions should be included in their community of interest, and the locations the commenter mentions should be excluded from their community of interest. Comment sentiment and clarity are both scaled to range from 0 to 1, with 0 representing negative sentiment and unclear comments while 1 represents positive sentiment and clear comments. The date of each comment was available from the original dataset. Ellmer was then used to estimate the complete set of Pennsylvania counties contained within each commenter's mentioned locations. Potential GeoNames entries were filtered to each set of counties, and the filtered list was converted into a registered tool used by Ellmer to reconcile mentioned locations with matches from the GeoNames database. 

```{r initialize-chat}

# initialize chat object ----
chat <- ellmer::chat_openai() # model gpt-4.1

```

```{r gather-comment-data}

# compile comment data ----
commentData <- purrr::map2(
  .x = as.character(tabulaPAHouse$Date[1:5]),
  .y = tabulaPAHouse$Description[1:5],
  .f = \(date, description) {
    
    ## gather individual comment data and mentioned locations ----
    chatOutput <- chat$chat_structured(
      description,
      type = ellmer::type_object(
        
        ### comment sentiment ----
        commentSentiment = ellmer::type_number(
          description = paste0(
            "Positive/Negative sentiment of this Pennsylvania commenter's description ",
            "of a community of interest scaled from 0 to 1. ",
            "A score of 0 indicates completely negative emotional sentiment, ",
            "while a score of 1 indicates completely positive emotional sentiment."
          )
        ),
        
        ### comment clarity ----
        commentClarity = ellmer::type_number(
          description = paste(
            "Clarity of this Pennsylvania commenter's description",
            "of a community of interest scaled from 0 to 1.",
            "Clarity should be based on how specific the commenter's mentioned",
            "locations are and how clearly they state their desire to include or",
            "exclude those locations from a community of interest.",
            "A score of 0 indicates no clarity,",
            "while a score of 1 indicates complete clarity."
          )
        ),
    
        ### locations included in the community of interest ----
        LocationsIncluded = ellmer::type_array(
          items = ellmer::type_string(),
          description = paste(
            "List of all locations that this Pennsylvania commenter mentions",
            "should be kept together in their community of interest.",
            "Should be exclusive and brief. "
          )
        ),
    
        ## locations excluded from the community of interest ----
        LocationsExcluded = ellmer::type_array(
          items = ellmer::type_string(),
          description = paste(
            "List of all locations that this Pennsylvania commenter mentions",
            "should be kept separate from their community of interest.",
            "Should be exclusive and brief. "
          )
        ),
        
        ## counties included in the community of interest ----
        ## TODO: improve fips code matching
        countiesIncluded = ellmer::type_array(
          items = ellmer::type_number(),
          description = paste(
            "Two or Three digit FIPS codes of all the counties in which this",
            "Pennsylvania commenter's mentioned locations might be contained."
          )
        )
      )
    )
    
    ## compile information for an individual comment ----
    commentInformation <- list(
      Date = date,
      OriginalComment = description,
      Characters = nchar(description),
      Sentiment = chatOutput$commentSentiment,
      Clarity = chatOutput$commentClarity,
      LocationsIncluded = chatOutput$LocationsIncluded,
      LocationsExcluded = chatOutput$LocationsExcluded,
      CountiesIncluded = chatOutput$countiesIncluded
    )
    
    ## return comment information ----
    return(commentInformation)
  }
)

```

```{r match-geonames}

# compile comment data ----
commentData <- commentData |> 
  purrr::map(
    .f = \(comment) {
      
      ## filter geonames matches to mentioned counties ----
      regionalGeoNames <- pennsylvaniaGeoNames |>
        dplyr::filter(Admin2Code %in% comment$CountiesIncluded)
      print(paste(nrow(regionalGeoNames), "possible GeoNames matches found"))
      
      ## register names filtering tool ----
      chat$register_tool(
        tool = ellmer::tool(
          name = "gather_geonames",
          description = "Gathers Pennsylvania GeoNames locations for possible matches",
          fun = \() return(regionalGeoNames$Name)
        )
      )
      
      ## add geonames matches included in community of interest ----
      Sys.sleep(time = 20)
      comment[["GeoNamesIncluded"]] <- chat$chat(
        paste(
          "Which GeoNames locations in Pennsylvania match these specific locations:",
          paste(comment$LocationsIncluded, collapse = ", "), "?",
          "Trim the response to a comma-separated location vector",
          "If no matches can be found, respond 'NA'"
        )
      ) |> as.character()
      
      ## add geonames matches excluded from community of interest ----
      Sys.sleep(time = 20)
      comment[["GeoNamesExcluded"]] <- chat$chat(
        paste(
          "Which GeoNames locations in Pennsylvania match these specific locations:",
          paste(comment$LocationsExcluded, collapse = ", "), "?",
          "Trim the response to a comma-separated location vector.",
          "If no matches can be found, respond 'NA'"
        )
      ) |> as.character()
      
      # return comment with added fields ----
      return(comment)
    }
  )

```

# References